\documentclass{article}

\usepackage[a4paper, margin=15mm]{geometry}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumitem}



\title{Homework 3 - Principles of Software Engineering (SWE311)}
\author{Alfaifi, Ammar}
\date{}

\begin{document}

\maketitle

\abstract{

As a team lead, I understand that my primary responsibility is to work in the best interest of the company. However, I also acknowledge that I have social and ethical responsibilities that may have consequences for the software development community and society at large.
}

\section{Advantages and Disadvantages of Using GPL Component}
In this scenario, using an open-source component distributed under GNU General Public License (GPL) has both advantages and disadvantages.

\subsection{Advantages}
\begin{itemize}
    \item It is cost-effective as it is free to use.
    \item It allows for the creation of high-quality software quickly and efficiently.
    \item It promotes collaboration among developers, which can lead to the creation of better software.
    \item It encourages innovation and creativity.
\end{itemize}

\subsection{Disadvantages}
\begin{itemize}
    \item Any software developed using a GPL component must be released as open source. This can make it difficult for companies to maintain a competitive edge.
    \item The company may lose control over the software development process, as others can modify and distribute the code.
\end{itemize}

\section{Course of Action}
In this situation, I would explain to the higher administration of the company the implications of using a GPL component and the licensing agreement that comes with it. I would discuss the advantages and disadvantages of using an open-source component and the potential consequences of not releasing the code as open source.

If the company still decides not to release the code as open source, I would respect their decision. However, I would also ensure that the code is not in violation of the GPL license. I would work with the development team to create a plan to remove the GPL component and replace it with a proprietary component. I would also make sure that the development team understands the importance of complying with licensing agreements and the potential consequences of violating them.

\section{Conclusion}
In conclusion, as a team lead, I have a responsibility to work in the best interest of the company, but I also have social and ethical responsibilities. In this scenario, I would prioritize the company's interest while ensuring that the team complies with the GPL license. I would also work to create a plan to replace the open-source component with a proprietary component to maintain the company's competitive edge.

\end{document}
